Nature or Nurture? Collisionless Evolution of Galactic Disc-Halo Systems

### 2019-07-19 Thesis/Paper_III_old.tex
Night before submission. Likely final pre-submission draft. Checked for spelling issues.

\newpage
\section{Abstract}

We present and study vertical structure evolution twelve high resolution simulations of Milky Way stellar discs inserted into cosmological dark matter haloes. The simulations exhibit a wide variety of vertical structure formation arising from bar buckling, interactions with the host halo, and substructure. A systematic study of the effects of the disc's initial dynamical temperature, measured by Toomre $Q$, and the time of insertion on the disc properties is conducted. Inserting a disc into a more equilibrated halo appears to lead to quieter vertical structure evolution, and increasing the disc's internal dynamical temperature appears to have a non-trivial impact. Furthermore, we identify a population of stars that get kicked up to high latitudes in each simulation. Our results are consistent with the idea that a disc-satellite interaction origin of these structures requires a perturber in the region of $10^{11} \, \solarm$ in close proximity to the disc. A broader interaction with the triaxial host halo is much better at generating these higher latitude structures on short timescales.  As such, we determine that a wide variety of cosmological environments are able to explain low-latitude disc structures like Monoceros, A13, and TriAnd.

\vspace{1in}

\section{Introduction} \label{sec:introduction}
%Literature review about vertical structure.

@@

@@

@@

Our paper is then organized as follows. To motivate our simulation suite, we present two toy models in \S\ref{sec:overview} to serve as discussion points. In particular, they focus on the important effect of inertial acceleration of the disc.  The time evolution of key structural statistics in all models is studied in \S \ref{sec:evolution}. Focusing on the outer disc, we examine how stars can be stripped to high latitudes in \S \ref{sec:kud}. To connect the previous results to their halo environments, we examine several representative evolution scenarios in more detail in \S\ref{sec:disentangle}. These scenarios include a disc-halo misalignment dominated system, a substructure interaction dominated system, and a system which is somewhere in between. Finally, we discuss our key results in \S\ref{sec:discussion}.

@@

In this section, we illustrate the factors driving vertical structure in stellar discs with two toy models.
\subsection{Quantifying Structure in Discs}
%\vspace{0.6in}

The analysis of simulated galaxies invariably involves computing averages of physical quantities, namely functions of the phase-space coordinates, $\textbf{r}$ and $\textbf{v}$. For some observable function,  $\Theta(\textbf{r},\textbf{v})$, the average in a region $S$ is,
\begin{eqnarray}
\langle\Theta\rangle_S &=&\frac{1}{M_S} \int_S \text{d}^3 \textbf{r} \text{d}^3\textbf{v} f(\textbf{r}, \textbf{v}) \Theta(\textbf{r}, \textbf{v})\\
&\approx& \frac{1}{M_S} \sum_{j \in S} m_j \Theta(\textbf{r}_j, \textbf{v}_j) \label{eq:z_statistic}
\end{eqnarray}
where $M_S$ is the mass in $S$, $f(\textbf{r},\textbf{v})$ is the phase space distribution function, and $m_j$ is the mass of the $j$-th particle in $S$. For stellar discs, we are often interested in the Fourier series representation of $\Theta$ to quantify $m$-fold angular symmetries in some annulus with mean radius, $R_S,$
\begin{equation}
\langle\Theta(R_S,m)\rangle_S = \frac{1}{M_S} \sum_{j \in S} e^{i m \phi_j} m_j \Theta(\textbf{r}, \textbf{v})
\end{equation}
where $m$ is the degree of symmetry, and $\phi_j$ is the azimuthal angle of particle $j$. When $\Theta = z^n$, we call the average $Z_m^n$. We make extensive use of this form for this paper.

@@

% This experiment is similar to Sellwood 1996...
The disc has an exponential surface density with scale length $R_d=3.7\,\kpc$ and mass $M_d=4\times10^{10} \,\solarm$. The halo is taken to be NFW, with scale length $R_h=33.2\,\kpc$, virial mass $M_h=1.1\times10^{12}$, and concentration $c=6.7$\footnote{The parameters for the halo are taken to be a sphericalized version of simulation B.Late in Table \ref{tab:sims}. This simulation will be described in more detail in \S\ref{sec:description}.}. We use $10^6$ halo particles and $3.5 \times 10^{6}$ disc particles. The subhalo is also assumed to be NFW with scale length $R_h=1.70\,\kpc$, virial mass $M_h=2.7\times10^{10}$, and concentration $c=38.1$\footnote{ The parameters for the substructure are given by B.Late.a in Table \ref{tab:substructure}.}. The subhalo is given a height of $z=100\,\kpc$ and zero velocity.

@@

The oscillation pattern seen in Fig. \ref{fig:toy_model_1_mean_height} is not stationary in time, but rather becomes more complicated as time goes on. Immediately after the displacement first changes signs, the pattern starts to propagate outwards more slowly as evidenced by decreasing (increasing magnitude) slope of the ridges. The mean height shows oscillating single fixed-point standing wave behaviour, at some points (around 1 Gyr) appearing to gain another node.  As the system equilibrates, the $m=0$ disturbances are pushed further to the outer disc, but the magnitude of the perturbations remains high.

@@

What is interesting about this experiment is the very long-lived nature of the flapping in the disc. Vertical instabilities are generally expected to phase mix, causing the disc to thicken. The flapping in our disc persists for longer than a dynamical time. In fact, a similar observation was made in an experiment performed by \citet{sellwood_1996} where a disc with similar properties was given an initial bulk displacement that varied as a function of radius. While other bending behaviour is rapidly suppressed, a flapping motion remained in the perturbed disc. Qualitatively, the results of \citet{sellwood_1996} are similar in magnitude to our own, suggesting that flapping naturally arises from near-axisymmetric perturbations of the disc.

@@

Here, we present a more complicated satellite encounter. The orbit is inspired by a satellite in one of our cosmological simulations, and shows a massive satellite on a close orbit with the disc, passing through the midplane several times. The parameters for the disc, halo, and subhalo are the same as in \S\ref{ssec:toy_model_1}.

We plot the $z$ coordinate of the subhalo, disc, and halo peak densities in Fig. \ref{fig:toy_model_2_orbits}.
Initially, the subhalo is far from the disc at over 100 kpc, and about 20 kpc above the disc in $z$. We see the disc and halo peak densities track each other almost exactly over the course of the simulation, consistent with Fig. \ref{fig:toy_model_1_mean_height} and Fig. \ref{fig:toy_model_1_mean_height_halo}. In some short frame of time, we suggest that this system can be viewed as a two body problem between the disc-halo system and subhalo. That is, we have motion described by,
\begin{equation}
\frac{M_{dh}\textbf{r}_{dh} +  + M_{sh} \textbf{r}_{sh}}{M_{dh} + M_{sh}} = \textbf{r}_{c,0} + \textbf{v}_{c,0} t
\end{equation}
where $M_{dh}$ is the implied two-body disc-halo mass, $M_{sh}$ is the implied two-body subhalo mass, $\textbf{r}_{dh} = (M_d \textbf{r}_d + M_h \textbf{r}_h)/M_{dh}$, $\textbf{r}_d$ is the disc peak density position, $\textbf{r}_h$ is the halo peak density position, $\textbf{r}_{sh}$ is the subhalo peak density position, $\textbf{r}_{c,0}$ is the initial implied center of mass, and $\textbf{v}_{c,0}$ is the implied center of mass velocity.

We can find the density modes from the particle distributions each time very easily. However, $M_{dh}$, $M_{sh}$, $\textbf{r}_{c,0}$, and $\textbf{v}_{c,0}$ are unknowns. We maximize the likelihood,
\begin{equation}
\ln \mathcal{L} = \left\vert \frac{M_{dh}\textbf{r}_{dh} +  + M_{sh} \textbf{r}_{sh}}{M_{dh} + M_{sh}} - \textbf{r}_{c,0} + \textbf{v}_{c,0} t \right\vert^2 \, \text{kpc}^{-2}
\end{equation}
in two 1.0 Gyr windows  centered on 0.5 Gyr and 2.0 Gyr, respectively. These windows are chosen since they encompass the time before pericentric passages at 0.85 Gyr and 3.0 Gyr, when the subhalo is losing the most mass. The idea is that during pericentric passages, we expect the system to be poorly described by a two-body fit. The two center of mass motion lines, $\textbf{r}_{c,0} + \textbf{v}_{c,0} t$ are shown as the solid black and dashed black lines in Fig. \ref{fig:toy_model_2_orbits}, respectively. We imply effective two-body mass ratios of 16.3:1 and 85.0:1, respectively, for disc-halo:subhalo masses.

The likelihood represents the mean distance deviation from our fit. For the two windows centered at 0.5 Gyr and 2.0 Gyr, we have the average root-log-likelihood $(N^{-1}\ln\mathcal{L})^{1/2}= 0.19$ in the first window and  $(N^{-1}\ln\mathcal{L})^{0.5} = 0.29$ in the second, where $N$ is the number of observations in the window. We thus infer small deviations from a two body problem when the subhalos are not at pericenter, implying that this model describes the data well.  We submit this technique as a way to quantify the effect of inertial acceleration in isolated satellite encounters. Although the drift velocity of the system changes as the satellite sheds mass, the picture is clear: the disc is placed on an oscillating orbit in $z$, and is no doubt subject to differential vertical acceleration. Despite this acceleration, the disc axis as measured by its moment of inertia tensor is relatively undeflected, being constrained to less than $2^{\circ}$ in the first 4 Gyr.

We see the orbit's effect on the mean height of the disc in Fig. \ref{fig:toy_model_2_mean_height}. At the beginning of the simulation, although the amplitude of the oscillations is lower, we see qualitatively similar behaviour to the first toy model. There is no obvious mean height bump when the substructure is at pericenter, indicating that passages through the disc plane are less substantial for generating $m=0$ corrugations than the overall two-body orbit. We expect that any subhalo capable of creating two-body motion of the disc along the disc axis will generate substantial, several hundred parsec corrugations.

@@

In this section, we proceed to generalize to a fully cosmological environment.

@@

For our pure dark matter simulation, we implement the zoom-in
technique of \citet{KatzQuasarZoom} and \citet{NavarroWhiteZoom}.
Selecting for Milky Way-like halos with no major mergers, we use the
results of \cite{onorbe_etal_2014} to ensure no low-resolution particle contamination.
We choose cosmological parameters based on the results
from Planck 2013 \citep{planck_2014} with $H_0=67.9\,{\rm
  km\,s}^{-1}\,{\rm kpc}^{-1}$, $\Omega_b = 0.0481$, $\Omega_0 =
0.306$, $\Omega_\Lambda = 0.694$, $\sigma_8 = 0.827$, and $n_s =
0.962$.  N-body initial conditions for the dark matter particles are
generated with the \textsc{music} code \citep{music}.

@@

%We apply the technique described in \S \ref{ssec:disc_insertio} to understand a possible disc origin of low-latitude streams, and more generally how vertical structure evolves in pure stellar discs. In particular, we are interested in understanding how, in a cosmological simulation, stars are kicked up out of the disc. They comprise a component, if not all member stars of low-latitude streams. We discuss useful definitions of this population subsequently in \S\ref{ssec:def}. Such kicked-up disc stars (KUD stars) may be useful probes of the Galactic potential if we can understand how their orbits were altered.

@@

Although we explore all of our models in \S\ref{sec:evolution} and \S\ref{sec:kud}, we identify three key simulations for further study in \S\ref{sec:disentangle}.
\begin{enumerate}
\item \textbf{Strong Subhalo Interaction:} B.Late has a close interaction with two subhaloes, but no substantial disc-halo misalignment. One of these subhaloes passes twice through the plane of the disc, one time interior to 30 kpc.
\item \textbf{Disc-Halo Misalignment:} C.Early presents a case with no massive subhalo interactions, but a substantial misalignment from its host halo.
\item \textbf{Distant Massive Subhaloes:} A.Early is relatively aligned to its host halo, but its most massive substructures do not interact closely with the disc.
\end{enumerate}
These simulations represent three archetypal halo environments, which we explore in more detail in subsequent sections.

@@

%As mentioned in \S\ref{ssec:numerical_experiments}, the haloes we consider are meant to resemble the Milky Way. In Table \ref{tab:sims}, all of the simulations we consider are listed.

%In addition to basic properties like halo mass, $M_h$, and scale radius, $R_h$, we also want to know how equilibrated the halo is with the disc. Discs will tumble in response to an external specific torque, $\boldsymbol \tau$, induced by the non-equilibrated halo. This invariably produces vertical structure.

Here, we further quantify our halo environments in terms of the effects we expect their smooth distributions and subhalos to have on the disc. For the former of these, we examine the following timescale,
\begin{equation}
T_\tau = \frac{L(R)}{\tau_\perp(R)},
\end{equation}
where $L(R)$ is the magnitude of the angular momentum in a ring centered at radius $R$, and $\tau_\perp(R)$ is the cylindrically averaged perpendicular torque in an annulus centered at $R$. This timescale is plotted as a function of radius at initialization for our models in Fig. \ref{fig:ratio_freqs}. This lower the timescale, the more the initial halo configuration contributes to the relevant torque. For example, we would interpret the simulation C.Early as being substantially non-equilibrated at initialization. Generally speaking, the more misaligned disc-halo systems have shorter torque action timescales. The effect of having a warmer disc is most pronounced in the inner regions, where the increased angular momentum from asymmetric drift appears to increase $T_\tau$.

%In a non-equilibrium state, shells of the halo may be misaligned from others at different radii. This creates a torquing effect on the disc and causes it to reorient (see \cite{DubinskiKuijkenRigidDisks}, for example). Torques present an opportunity for the angular momenta of orbits to change, and may result in the generation of KUD stars. Fig. \ref{fig:disk_torques} shows the specific torque as a function of radius for all models. In the most misaligned simulations, bauer2018b.Early/Warm, bauer2018b.Late, and C.Early/Warm, there is an obvious torque signature in the intermediate parts of the disc.

%he relative tilt of the halo principal axes compared to the halo interior for all runs. The model considered in \citet{bauer2018b}, bauer2018b.Early stands out the most. It shows substantial misalignment at all radii. C.Early also shows substantial misalignment, but the other simulations seem to have mostly aligned short axes.\

@@

\begin{figure}
    \centering
	\includegraphics[width=0.8\textwidth]{../figures/surface_density_ratios.eps} \caption{The ratio of surface density, $\Sigma$, to azimuthally averaged surface density, $\Sigma_0$ for the C.Early, C.Warm, and C.Late models. The circles correspond to $R_p = 2.2\, R_d = 8.1 \,\kpc$ and $20 \,\kpch = 29.5 \, \kpc$.}
	\label{fig:final_densities}
\end{figure}

@@

In isolation, a disc-halo system forms a bar when shot noise induced perturbations exponentially grow \citep{EfstathiouShotNoise}. Factors such as the relative contribution of the disc to the rotation curve, radial velocity dispersion, and disc scale height broadly determine bar formation rate, strength, and length \citep{AthanassoulaSellwood1986,ChristodoulouStability1995, Klypin2009, Sellwood2013, bauer2018b}. 
Strong environmental perturbers such as halo substructure and tides can induce bar formation even in discs that resist bar formation in isolation \citep{gauthier_2006, kazantzidis2008, purcell2011}. In similar studies using simpler disc insertion schemes, \citet{debuhr_2012,ys_2015} used a disc insertion scheme, and found that stellar discs always form bars in a cosmological environment, even if they are submaximal. \citet{ys_2015} found that the presence of a classical bulge substantially mitigated this conclusion.

In a previous paper, we argued that these simulations overestimated the susceptibility of the disc to forming strong bars \citep{bauer2018b}. This argument stems from the initial thickness used in these studies, which we found to be a major factor in determining present-day bar strength. Specifically, this is because thin discs grow bars more quickly, and are able to buckle more easily, a process which regulates their strength. In contrast, thick discs grow their bars almost completely monotonically.

The final, present-day surface density excesses are shown in Fig. \ref{fig:final_densities}. Broadly speaking, our simulations form bars similar to the Milky Way's bar.  The bulk of the bar mass in all simulations appears contained to within $R_p = 2.2\, R_d$, as shown by visual inspection of the inner circles. Furthermore, it the warm suite is not substantially distinguishable upon inspection of this snapshot, indicating that the increased radial velocity dispersion did little to stop the formation of a bar in a cosmological setting.   The late-insertion suite shows a reduced bar length, bolstering an idea presented in \citet{bauer2018b} that for similar galaxies, one might be able to use the bar length to compare relative dates of stellar discs. Although we show only the the C.Early, C.Warm, and C.Late models, we find no substantial differences with other models within each subsuite.

In Fig. \ref{fig:a_2}, we plot the bar strength parameter (Eq. \eqref{eq:z_statistic} with $\theta=1$ and $m=2$)  evaluated in the region $R < R_p$ \citep{debattista_sellwood_2000}. First, we note that all models experience some kind of buckling event. We notice that the late suite appears to buckle lower magnitudes. Generally speaking, we see that the warm discs are delayed in their bar formation and buckling, although bauer2018b.Warm is an exception. The final bar strengths as measured by $A_2(R<R_p)$ are similar in magnitude, consistent with Fig. \ref{fig:final_densities}.

The picture painted between Fig. \ref{fig:final_densities} and Fig. \ref{fig:a_2} bolsters the previous claim that Milky Way discs in a cosmological environment will all form bars. Upon inspection, the bars are qualitatively similar to Milky Way's bar, suggesting no real problem on this front for a Milky Way thin stellar disc forming around or slightly after $z=1$. If the Milky Way took a similar trajectory to these simulations, it would not be unreasonable to suspect that observed vertical structure in the Milky Way is in part driven by a buckling bar.

@@

In fact, we see this confirmed in our simulations.  Fig. \ref{fig:z_rms} shows the evolution of $z_{rms}$ for each disc in the fiducial and warm suites. We see that substantial heating follows bar buckling in each simulation, a result consistent with \citet{ys_2015,bauer2018b}. We note that the warm suite exhibits up to 25\% more disc heating.  There is very little to say about the late-insertion scheme in this respect other than that the discs approximately double in height over the simulation time. These simulations end as the buckling process begins, and this may explain why vertical heating is more limited.\\

@@

@@

%\begin{equation}
%Z_m(S) = \left\vert \frac{1}{M_S} \sum_{j \in S} m_j z_j e^{i m \phi_j} \right\vert, \label{eq:z1}
%\end{equation}
%where $S$,$M_S$, and $\phi_j$ are defined as in Eq. \ref{eq:a2}, $z_j$ is the displacement of the $j$-th particle in $S$, and $m=1$ for the relevant bending mode. We take $S$ to be the annulus, $R \in [4.0 \,R_d,4.5\,R_d]$. The bending mode is plotted against time for all of our simulations in Fig. \ref{eq:z1}.  Milky Way discs are known to spontaneously form bending waves in isolation due to shot noise \citep{chequers_2017}, as well as in response to realistic populations of subhaloes \citep{chequers_2018}. In the absence of a major perturber, as in \citet{chequers_2018}, the $m=1$ height bending mode is constrained to less than but on the order of $10^{-1}$ in the region we are considering.

>>>>

\newpage
\section{Abstract}
We present and study vertical structure evolution twelve high resolution simulations of Milky Way stellar discs inserted into cosmological dark matter haloes. The simulations exhibit a wide variety of vertical structure formation arising from bar buckling, interactions with the host halo, and substructure. A systematic study of the effects of the disc's initial dynamical temperature, measured by Toomre $Q$, and the time of insertion on the disc properties is conducted. Inserting a disc into a more equilibrated halo appears to lead to quieter vertical structure evolution, and increasing the disc's internal dynamical temperature appears to have a non-trivial impact. Furthermore, we identify a population of stars that get kicked up to high latitudes in each simulation. Our results are consistent with the idea that a disc-satellite interaction origin of these structures requires a perturber in the region of $10^{11} \, \solarm$ in close proximity to the disc. A broader interaction with the triaxial host halo is much better at generating these higher latitude structures on short timescales.  As such, we determine that a wide variety of cosmological environments are able to explain low-latitude disc structures like Monoceros, A13, and TriAnd. 
\section{Introduction} \label{sec:introduction}
%Literature review about vertical structure.

@@

%Here, bar buckling is proposed to have a long-lasting impact on the vertical structure of a stellar disc. Since bars buckle in response to unstable bending waves, it is possible that the effect of bar buckling in the Milky Way is long lived and observable in the thin disc.

%

@@

Our paper is then organized as follows. To motivate our simulation suite, we present two toy models in \S\ref{sec:overview} to serve as discussion points. In particular, they focus on the important effect of inertial acceleration of the disc. We then describe twelve simulations of discs inserted into cosmological haloes in \S \ref{sec:description}. Although the halo evironments have similar total mass to the Milky Way, they vary widely in terms of their substructure and smooth density components. The time evolution of key structural statistics in all models is studied in \S \ref{sec:evolution}. Focusing on the outer disc, we examine how stars can be stripped to high latitudes in \S \ref{sec:kud}. To connect the previous results to their halo environments, we examine several representative evolution scenarios in more detail in \S\ref{sec:disentangle}. These scenarios include a disc-halo misalignment dominated system, a substructure interaction dominated system, and a system which is somewhere in between. Finally, we discuss our key results in \S\ref{sec:discussion}.

@@

@@

In this section, we illustrate the factors driving vertical structure in stellar discs with two toy models.
\subsection{Quantifying Structure in Discs}
%\vspace{0.6in}
We choose to approach summarizing our models in this paper by looking at statistics of the form,
\begin{equation}
Z_m^n  = \left \vert \frac{1}{M_s} \sum_{j \in S} m_j e^{i m \phi_h} z^n_j \right\vert \label{eq:z_statistic}
\end{equation}
where $M_S$ is the mass in some region $S$, $m_j$ is the mass of the $j$-th particle in $S$, and $\phi_j$ is the azimuthal angle of particle $j$, and $z_j$ is its height. The quantities $m$ and $n$ are chosen to give the statistic a desired meaning. For example, $(m,n) = (0,1)$ is the mean height in $S$, whereas $(m,n) = (0,2)$ is the mean-squared height. Higher values of $m$ will pick out certain symmetries. In the case of $m=1$, ring-like vertical displacements show up. The special case of $(m,n) = (2,0)$ will be high if there are planar features with two symmetry axes, like bars. These special cases of Eq. \eqref{eq:z_statistic} have important significance in the evolution of spiral galaxies, as we discuss throughout the paper.
\\

@@

% This experiment is similar to Sellwood 1996...
The disc has an exponential surfacec density with scale length $R_d=3.7\,\kpc$ and mass $M_d=4\times10^{10} \,\solarm$. The halo is taken to be NFW, with scale length $R_h=33.2\,\kpc$, virial mass $M_h=1.1\times10^{12}$, and concentration $c=6.7$\footnote{The parameters for the halo are taken to be a sphericalized version of simulation B.Late in Table \ref{tab:sims}. This simulation will be described in more detail in \S\ref{sec:description}.}. We use $10^5$ halo particles and $3.5 \times 10^{5}$ disc particles. The subhalo is also assumed to be NFW with scale length $R_h=1.70\,\kpc$, virial mass $M_h=2.7\times10^{10}$, and concentration $c=38.1$\footnote{ The parameters for the substructure are given by B.Late.a in Table \ref{tab:substructure}.}. The subhalo is given a height of $z=100\,\kpc$ and zero velocity.

@@

The oscillation pattern seen in Fig. \ref{fig:toy_model_1_mean_height} is not stationary in time, but rather becomes more complicated as time goes on. Immediately after the first sign reversal, the pattern starts to propagate outwards more slowly as evidenced by decreasing (increasing magnitude) slope of the ridges. The mean height shows oscillating single fixed-point standing wave behaviour, at some points (around 1 Gyr) appearing to gain another node.  As the system equilibrates, the $m=0$ disturbances are pushed further to the outer disc, but the magnitude of the perturbations remains high.

@@

What is interesting about this experiment is the very long-lived nature of the flapping in the disc. Vertical instabilities are generally expected to phase mix, causing the disc to thicken. The flapping in our disc persists for much longer than a dynamical time. In fact, a similar observation was made in an experiment performed by \citet{sellwood_1996} where a disc with similar properties was given an initial bulk displacement that varied as a function of radius. While other bending behaviour is rapidly suppressed, a flapping motion remained in the perturbed disc. Qualitatively, the results of \citet{sellwood_1996} are similar in magnitude to our own, suggesting that flapping naturally arises from near-axisymmetric perturbations of the disc.

@@

Here, we present a more complicated satellite encounter. The orbit is inspired by a satellite in one of our cosmological simulations, and shows a massive satellite on a close, polar orbit with the disc. The parameters for the disc, halo, and subhalo are the same as in \S\ref{ssec:toy_model_1}.

We plot the $z$ coordinate of the subhalo, disc, and halo peak densities in Fig. \ref{fig:toy_model_2_orbits}.
Initially, the subhalo is far from the disc at around 100 kpc, and about 50 kpc above the disc in $z$. We see the disc and halo peak densities track each other almost exactly over the course of the simulation, consistent with Fig. \ref{fig:toy_model_1_mean_height} and Fig. \ref{fig:toy_model_1_mean_height_halo}. In some short frame of time, we suggest that this system can be viewed as a two body problem between the disc-halo system and subhalo. That is, we have motion described by,
\begin{equation}
\frac{M_{dh}\textbf{r}_{dh} +  + M_{sh} \textbf{r}_{sh}}{M_{dh} + M_{sh}} = \textbf{r}_{c,0} + \textbf{v}_{c,0} t
\end{equation}
where $M_{dh}$ is the implied two-body disc-halo mass, $M_{sh}$ is the implied two-body subhalo mass, $\textbf{r}_{dh} = (M_d \textbf{r}_d + M_h \textbf{r}_h)/M_{dh}$, $\textbf{r}_d$ is the disc peak density position, $\textbf{r}_h$ is the halo peak density position, $\textbf{r}_{sh}$ is the subhalo peak density position, $\textbf{r}_{c,0}$ is the initial implied center of mass, and $\textbf{v}_{c,0}$ is the implied center of mass velocity.

We can find the density modes from the particle distributions each time very easily. However, $M_{dh}$, $M_{sh}$, $\textbf{r}_{c,0}$, and $\textbf{v}_{c,0}$ are unknowns. We maximize the likelihood,
\begin{equation}
\ln \mathcal{L} = \left\vert \frac{M_{dh}\textbf{r}_{dh} +  + M_{sh} \textbf{r}_{sh}}{M_{dh} + M_{sh}} - \textbf{r}_{c,0} + \textbf{v}_{c,0} t \right\vert^2
\end{equation}
in two 1.0 Gyr windows centered on the pericenters at 0.5 Gyr and 2.0 Gyr. The two center of mass motion lines, $\textbf{r}_{c,0} + \textbf{v}_{c,0} t$ are shown as the solid black and dashed black lines in Fig. \ref{fig:toy_model_2_orbits}, respectively. At first and second pericenters, we imply effective two-body mass ratios of 14.4:1 and 29.2:1, respectively, for disc-halo:subhalo masses.

The first pericenter mass ratio is in good agreement with the distance to center of mass ratio of the disc-halo and subhalo, which is 13.2:1. The fact that this is close to the mass ratio is a good indication that two body motion describes the disc-halo system's response to the subhalo in the time around the first pericentric passage. This is because the distance ratio was not a parameter in our fit, but we know it must be the same as the mass ratio for true two-body motion. We submit this technique as a way to quantify the effect of inertial acceleration in isolated satellite encounters.

Although the drift velocity of the system changes as the satellite sheds mass, the picture is clear: the disc is placed on an oscillating orbit in $z$, and is no doubt subject to differential vertical acceleration. 
We see the orbit's effect on the mean height of the disc in Fig. \ref{fig:toy_model_2_mean_height}. At the beginning of the simulation, although the amplitude of the oscillations is lower, we see qualitatively similar behaviour to the first toy model. There is no obvious mean height bump when the substructure is at pericenter, indicating that passages through the disc plane are less substantial for generating corrugations than the overall orbit. We also note that the disc orientation is relatively undeflected during this time period, being constrained to less than $5^{\circ}$ in the first 2.5 Gyr.

@@

With the findings of \S\ref{sec:overview} in mind, we proceed to generalize to a fully cosmological environment. In this section, we give an overview of our cosmological simulations. We begin by summarizing the disc insertion technique. Then, we give a detailed description of the individual halo environments. Finally, we present broad results on the evolution of bars and vertical structure in our models.

@@

For our pure dark matter simulation, we implement the zoom-in
technique of \citet{KatzQuasarZoom} and \citet{NavarroWhiteZoom}.
Selecting for Milky Way-like halos with no major mergers, we use the
results of \cite{onorbe_etal_2014} to ensure no low-resolution particle contamination 
We choose cosmological parameters based on the results
from Planck 2013 \citep{planck_2014} with $H_0=67.9\,{\rm
  km\,s}^{-1}\,{\rm kpc}^{-1}$, $\Omega_b = 0.0481$, $\Omega_0 =
0.306$, $\Omega_\Lambda = 0.694$, $\sigma_8 = 0.827$, and $n_s =
0.962$.  N-body initial conditions for the dark matter particles are
generated with the \textsc{music} code \citep{music}.

@@

@@

Although we explore all of our models in \S\ref{sec:evolution} and \S\ref{sec:kud}, we identify three key simulations for further study in \S\ref{sec:disentangle}.
\begin{enumerate}
\item \textbf{Strong Subhalo Interaction:} B.Late has a close interaction with two subhaloes, but no substantial disc-halo misalignment. One of these subhaloes passes twice through the plane of the disc, one time interior to 30 kpc.
\item \textbf{Disc-Halo Misalignment:} C.Early presents a case with no massive subhalo interactions, but a substantial misalignment from its host halo.
\item \textbf{Distant Massive Subhaloes:} A.Early is relatively aligned to its host halo, but its most massive substructures do not interact closely with the disc.
\end{enumerate}
These simulations represent three archetypal halo environments, which we explore in more detail in the next section.

@@

Here, we further quantify our halo environments in terms of the effects we expect their smooth distributions and subhalos to have on the disc. For the former of these, we examine the following timescale,
\begin{equation}
T_\tau = \frac{L(R)}{\tau_\perp(R)},
\end{equation}
where $L(R)$ is the magnitude of the angular momentum in a ring centered at radius $R$, and $\tau_\perp(R)$ is the cylindrically averaged perpendicular torque in an annulus centered at $R$. This timescale is plotted as a function of radius at initialization for our models in Fig. \ref{fig:ratio_freqs}. This lower the timescale, the more the initial halo configuration contributes to the relevant torque. For example, we would interpret the simulation C.Early as being substantially non-equilibrated at initialization. Generally speaking, the more misaligned disc-halo systems have shorter torque action timescales. The effect of having a warmer disc is most pronounced in the inner regions, where the increased angular momentum appears to increase $T_\tau$.

@@

\begin{figure}
    \centering
	\includegraphics[width=0.8\textwidth]{../figures/three_panel_densities.eps} \caption{The final face-on density projections for the C.Early, C.Warm, and C.Late models. The circles correspond to $R_p = 2.2\, R_d = 8.1 \,\kpc$ and $20 \kpch = 29.5 \, \kpc$.}
	\label{fig:final_densities}
\end{figure}

@@

In isolation, a disc-halo system forms a bar when shot noise induced perturbations exponentially grow \citep{EfstathiouShotNoise}. Factors such as the relative contribution of the disc to the rotation curve, radial velocity dispersion, and disc scale height broadly determine how a disc forms a bar in isolation \citep{AthanassoulaSellwood1986,ChristodoulouStability1995, Klypin2009, Sellwood2013, bauer2018b}.

In the presence of strong environmental perturbers, $m=2$ symmetric perturbations can induce rapid bar formation \citep{gauthier_2006, kazantzidis2008, purcell2011, bauer2018b}. In similar studies using simpler disc insertion schemes, \citet{debuhr_2012,ys_2015} found that pure stellar discs invariably formed large bars in cosmological environments. Although \citet{bauer2018b} found that disc-halo systems form bars, they found that the strength of the bar was not doomed to be unrealistic. Since \citet{bauer2018b} considered only one halo, the previously stated result is tenuous.

A full study of how bars form in our simulations is outside the scope of this work, but since \citet{bar_buckling_echo} suggests that the buckling of a bar is a significant driver of vertical structure, it is important to understand the bar strength history in our simulations. In particular, we are looking for variations in the buckling times of our discs.

%With the twelve simulations presented in this paper, we can complete our contribution to how pure stellar discs behave in a cosmological setting.
%\\

Broadly speaking, our simulations form bars similar to the Milky Way;s bar. The final, present-day snapshot for several models is shown in Fig. \ref{fig:final_densities}. The bulk of the bar mass in all simulations appears contained to within $R_p = 2.2\, R_d$, as shown by visual inspection of the inner circles. Furthermore, it is worth noting that the warm suite is not substantially distinguishable upon inspection of this snapshot, indicating that the increased radial velocity dispersion did little to stop the formation of a bar in a cosmological setting.   The late-insertion suite shows a reduced bar length, bolstering an idea presented in \citet{bauer2018b} that for similar galaxies, one might be able to use the bar length to compare relative dates of stellar discs. Although we show only the the C.Early, C.Warm, and C.Late models, we find no substantial differences with other models within each subsuite.

We can see the full time evolution of the bar strength in Fig. \ref{fig:a_2} as measured by the $A_2$ parameter. Conventionally, we define $A_m = Z^0_m$ \citep[for instance]{debattista_sellwood_2000}. We can evaluate $A_2$ in the region interior to $R_p$ as a global measure of bar strength. The top panel of Fig. \ref{fig:a_2} shows the evolution of $A_2(<R_p)$ for our fiducial and warm suites. Generally speaking, the increased vertical temperature has little-to-no effect on bar strength. The discs show increased bar growth until $a\sim0.6$, where the bars buckle and then resume growing for the remainder of the simulation. One notable exception is a slight difference prior to $a=0.75$ in the evolution of A.Early and A.Warm. In A.Warm, the buckling time is shifted to slightly after $a=0.7$, a result we might be able to understand in terms of the halo environment. From Table \ref{tab:sims}, we see that these simulations are characterized by very low disc-halo misalignment, and from Fig. \ref{fig:substructure_orbits} we see the most massive subhalo is quite far from the disc and on a tangential orbit. This relatively unperturbed state may be why the disc nurture matters more at early times. However, both models eventually end up quite similar at late times ($a>0.75$). The late-insertion suite shown in the bottom panels is similar to the top panel if it were shifted. The bar grows for a similar amount of time, and we can see the beginnings of buckling near the end of these simulations.
\\

The picture painted between Fig. \ref{fig:final_densities} and Fig. \ref{fig:a_2} suggests that Milky Way discs in a cosmological environment will all form bars, although these bars are far more realistic than the ones presented in previous studies using disc insertion techniques. Upon inspection, the bars are qualitatively similar to Milky Way's bar, suggesting no real problem on this front for a Milky Way thin stellar disc forming around or slightly after $z=1$. If the Milky Way took a similar trajectory to these simulations, it would not be unreasonable to suspect that observed vertical structure in the Milky Way is in part driven by a buckling bar.

@@

In fact, we see this confirmed in our simulations.  Fig. \ref{fig:z_rms} shows the evolution of the root-mean-squared height for each disc in the fiducial and warm suites. We see that substantial heating follows bar buckling in each simulation, a result consistent with \citet{ys_2015,bauer2018b}. An interesting point to make is that the warm suite exhibits up to 25\% more disc heating depending on the point in disc evolution. We understand this as being due to a rotation of local velocity ellipsoids out of the plane of the disc. When this happens, a larger initial $Q$ translates to a larger out-of-plane component, which would enhance vertical heating. In other words, increased radial dispersion amplifies secular heating mechanisms. There is very little to say about the late-insertion scheme in this respect other than that the discs approximately double in height over the simulation time. These simulations end as the buckling process begins, and this may explain why vertical heating is more limited.\\

@@

Bulk displacements beyond the Solar neighborhood could explain structures like the Monoceros Ring discovered by \citet{newberg_2002}. One thing we are particularly interested in is how the disk is bending in these regions as a function of time in the simulation. To this end, we want to quantify bending modes in the vicinity of where a Monoceros analogue would be. We look at the mean height and the $m=1$ bending mode, $Z_1 = Z_1^1$.

@@

\section{KUD Dynamics} \label{sec:kud}

The last section focused on bulk bending of the disc around the solar neighborhood. Here, we focus on stars which may be stripped from the disc to very high latitudes. We give a definition for this population and succinctly summarize its dynamics.
